Uni- and Multivariate Analysis of Dendritic Ca2+ Data

### 2017-07-17 presentation/demo.tex
asd

@@

\begin{frame}[fragile]{The Experiment}
\begin{columns}[T,onlytextwidth]
    \column{0.5\textwidth}
    	\only<+->{
      \begin{figure}
      \includegraphics[scale=0.1]{mouse.png}
      \end{figure}}

@@

\begin{frame}[fragile]{The Data - Neuronal}
	\begin{center}
	\begin{figure}
	\caption*{Trial-averaged $\text{Ca}^{2+}$ fluorescence traces of random dendrites}
      \includegraphics[scale=0.35]{traces.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{The Data - Behavioral}
	\begin{center}
	\begin{figure}
      \includegraphics[scale=0.23	]{lickplots.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{The Data - Psychometric Curve}
\begin{center}
	\begin{figure}
      \includegraphics[scale=0.38]{psychometric.png}
	\end{figure}
	\end{center}

@@

\section{Goal}
\begin{frame}[fragile]{Goal}
The goal of this project is to investigate the following:
\begin{itemize}
\only<+->{
\item What information can we extract from the data?}
\only<+->{
\item What can we gain by doing multivariate analysis as opposed to univariate?}
\only<+->{
\item Can we say something about population coding?}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{Stimulus Presence Detection}
\begin{itemize}
\item Feature: mean activity of dendrite over one second after stimulus
\item Class one: all trials with stimulus strength 0
\item Class two: all trials with selected stimulus strength $(\neq 0)$
\end{itemize}
\end{frame}

\begin{frame}[fragile]{ROC}
\begin{center}
	\begin{figure}
      \includegraphics[width=0.75\textwidth]{roc_max.png}
	\end{figure}
	\end{center}
\end{frame}

@@

@@

@@

\begin{frame}[fragile]{DI and SVM Dendrtie Rank Order Correlation}
\begin{center}
	\begin{figure}
	\caption*{Something is wrong here, should be more correlation}
      \includegraphics[width=1.0\textwidth]{DI_vs_SVM.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Tuning Curves}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{tuning_ca2+.png}
      \caption*{todo:accuracies}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Behavoreal Prediction - Most Accurate Dendrites}
\begin{table}
    \caption*{Stimulus strength 1 (near thereshold)}
    \begin{tabular}{c|c|c}
      \toprule
      Dendrite \# & Mean Accuracy & Standard Deviation\\
      \midrule
      70 & 0.78 & +/- 0.28\\
      47 & 0.77 & +/- 0.32\\
      56 & 0.73 & +/- 0.24\\
      92 & 0.70 & +/- 0.33\\
      \bottomrule
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}[fragile]{Behavoreal Tuning}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{tuning_ca2+_bb.png}
      \caption*{todo:accuracies}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Rank Correlations of Behavoreal and Presence Dendrites}
\begin{center}
	\begin{figure}
	\caption*{reduce ndends}
      \includegraphics[width=1.0\textwidth]{transrank.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{SVM Performance on Combined Dendrites}
Placeholder - Something wrong when using other files!
\end{frame}

\begin{frame}[fragile]{SVM Performance on Combined Dendrites - Behavoreal}
Placeholder - Something wrong when using other files!
\end{frame}

\begin{frame}[fragile]{Optimal Averaging Times}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=1.0\textwidth]{placeholder_wind_pres.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Optimal Averaging Times - Behavior}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=1.0\textwidth]{placeholder_wind_hit.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Mante - Approach (Placeholder Title)}
Math goes here
\end{frame}

\begin{frame}[fragile]{Mante - Approach - Results}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=1.0\textwidth]{reg_norm_placeholder.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Mante - Approach - Results}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=1.0\textwidth]{reg_inte_placeholder.png}
	\end{figure}
	\end{center}
\end{frame}

>>>>

\usepackage{bibentry}

@@

\begin{frame}[fragile]{The Experiment}
Takahashi et al. 2016	
\begin{columns}[T,onlytextwidth]
    \column{0.5\textwidth}
    	\only<+->{
      \begin{figure}
      \includegraphics[scale=0.1]{mouse.png}
      \end{figure}}

@@

\begin{frame}[fragile]{The Data - Neuronal}
	\begin{center}
	\begin{figure}
	\caption*{Trial-averaged $\text{Ca}^{2+}$ fluorescence traces of random dendrites}
      \includegraphics[width=\textwidth]{traces.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{The Data - Neuronal}
	\begin{center}
	\begin{figure}
	\caption*{Trial-averaged $\text{Ca}^{2+}$ fluorescence traces of random dendrites}
      \includegraphics[width=\textwidth]{traces_roi.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{The Data - Behavioral}
	\begin{center}
	\begin{figure}
      \includegraphics[width=\textwidth]{lickplot_single.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{The Data - Psychometric Curve}
\begin{center}
	\begin{figure}
      \includegraphics[width=\textwidth]{psychometric.png}
	\end{figure}
	\end{center}

@@

\section{Goal}
\begin{frame}[fragile]{Goal}
The goal of this project is to investigate the following:
\begin{itemize}
\only<+->{
\item What is the relationship between Ca2+ activity and stimulus intensity/behavior}
\only<+->{
\item Does a multivariate approach give us any significant advantage over a univariate one?}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]{Stimulus Presence Detection}
%\begin{itemize}
%\item Feature: mean activity of dendrite over one second after stimulus
%\item Class one: all trials with stimulus strength 0
%\item Class two: all trials with selected stimulus strength $(\neq 0)$
%\end{itemize}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{data_p.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{ROC}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{roc_exp.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{ROC - Best ROC Curves}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{roc_four.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{SVM - What is an SMV?}
\begin{center}
\only<+->{
\begin{figure}
      \includegraphics[height=0.3\textwidth]{svm_exp1.png}
	\end{figure}}
	\only<+->{
	\begin{figure}
      \includegraphics[height=0.3\textwidth]{svm_exp.png}
	\end{figure}}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Why SVM?}

\begin{itemize}
\item staple of machine learning
\item regularization parameter limits overfitting
\item good for small samples with many features (in theory)
\end{itemize}

@@

\begin{frame}[fragile]{Rank Order Correlation}
\only<+->{
We want to quantify the similarity between two rank orders of dendrites $R$ and $Q$, which both have length $n$.}

\only<+->{
Standard approach: Spearman's rank order coefficient 
$$
\rho(R,Q) = 1 - \frac{6\sum_i^nD_i^2}{n(n	^2-1)}
$$
where $D_i=R_i-Q_i$.}

\only<+->{
Problem: Many of the dendrites are not significant for classification but affect $\rho$ greatly. We would like to be able to \textbf{weight our} $\mathbf{D_i}$'s.}
\end{frame}

\begin{frame}{Weighted Rank Order Correlation Coefficient}
One possible soulution:
$$
\rho_W(R,Q) = 1-\frac{-2\sum_i^nw_iD_i^2}{\sum_i^nw_i(n-2i+1)^2}
$$
With a monotonicity constraint on $W$.
\end{frame}

@@

@@

\begin{frame}[fragile]{Tuning Curves}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{tuning_ca2+single.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Tuning Curves}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{tuning_ca2+.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Bevahior Detection}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{data_hm.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Behavioral Prediction - Most Accurate Dendrites}
\begin{table}
    \caption*{Different dataset - Stimulus strength 1 (near thereshold)}
    \begin{tabular}{c|c|c}
      \toprule
      Dendrite \# & $\mu_{acc}$ & $\sigma_{acc}$\\
      \midrule
      88 & 0.80 & 0.08\\
      57 & 0.80 & 0.08\\
      92 & 0.78 & 0.06\\
      56 & 0.76 & 0.10\\
      \bottomrule
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}[fragile]{Behavorial Predition - Imbalance}
\begin{center}
	\begin{figure}
	  \caption*{$N_{no-lick}$ vs. $N_{lick}$ for all stimuli}
      \includegraphics[width=1.0\textwidth]{fraction.png}
	\end{figure}
	\end{center}
\end{frame}

%\begin{frame}[fragile]{Behavoreal Tuning}
%\begin{center}
%	\begin{figure}
%      \includegraphics[width=1.0\textwidth]{tuning_ca2+_bb.png}
%      \caption*{todo:accuracies}
%	\end{figure}
%	\end{center}
%\end{frame}

\begin{frame}[fragile]{Rank Correlations of Behavoreal and Presence Dendrites}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{transrank.png}
	\end{figure}
	\end{center}
\end{frame}

@@

\begin{frame}[fragile]{Presence Detection - Feature Selection}
\only<+->{
Since we have only few datapoints per condition (30-50 per class) in comparison to the number of features (100-150), we are prone to overfit to noise.}

\only<+->{
$\Rightarrow$ we have to do \textbf{feature selection}.}

\only<+->{
We can use the previously best discriminating dendrites (SVM/DI).}
\end{frame}

\begin{frame}[fragile]{DI and SVM Dendrite Rank Order Correlation}
\begin{center}
	\begin{figure}
	%\caption*{todo: tens, improve weights}
      \includegraphics[width=1.0\textwidth]{rank_DI_SVM.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{SVM Performance on Combined Dendrites - Presence Detection}
\begin{center}
	\begin{figure}
	%\caption*{todo: tens, improve weights}
      \includegraphics[width=1.0\textwidth]{combined_presence.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{SVM Performance on Combined Dendrites - Presence Detection}
\begin{center}
	\begin{figure}
	%\caption*{todo: tens, improve weights}
      \includegraphics[width=1.0\textwidth]{combined_presence_alt.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{SVM Performance on Combined Dendrites - Behavioral}
\begin{center}
	\begin{figure}
	%\caption*{todo: tens, improve weights}
      \includegraphics[width=1.0\textwidth]{combined_hitmiss.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{SVM Performance on Combined Dendrites - Behavioral}
\begin{center}
	\begin{figure}
	%\caption*{todo: tens, improve weights}
      \includegraphics[width=1.0\textwidth]{combined_hitmiss_alt.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Optimal Averaging Times}
\begin{center}
	\begin{figure}
	\caption*{Dummy}
      \includegraphics[width=0.6\textwidth]{placeholder_wind_pres.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Optimal Averaging Times - Behavior}
\begin{center}
	\begin{figure}
	\caption*{Dummy}
      \includegraphics[width=0.6\textwidth]{placeholder_wind_hit.png}
	\end{figure}
	\end{center}
\end{frame}

\section{Population Coding}

\begin{frame}[fragile]{Population Coding - Idea (Mante, Sussillo 2013)}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{coord.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Population Coding - Idea (Mante, Sussillo 2013)}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{traj.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Population Coding - Idea (Mante, Sussillo 2013)}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{new_coord.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Population Coding - Idea (Mante, Sussillo 2013)}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{trafo.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Population Coding - Idea (Mante, Sussillo 2013)}
\begin{center}
	\begin{figure}
      \includegraphics[width=1.0\textwidth]{trafo_final.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=0.7\textwidth]{reg.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\begin{center}
	\begin{figure}
	\caption*{Solve Problem with combined SVM first}
      \includegraphics[width=0.7\textwidth]{reg_sum.png}
	\end{figure}
	\end{center}
\end{frame}

\begin{frame}
\begin{thebibliography}{9}
\bibitem {taka} Takahashi, N., Oertner, G. T., Hegemann, P. \& Larkum, E. M. Active cortical dendrites modulate perception. Science 335, 1587-1590 (2016)
\bibitem {mante} Mante V., Sussillo D., Shenoy KV., Newsome WT. Context-dependent computation by recurrent dynamics in prefrontal cortex. Nature 503, 78-84 (2013)
\bibitem {pinto} Pinto da Costa, J. New Results in Weighted Correlation and Weighted Principal Component Analysis with Applications, Chapter 2 (2015)
\end{thebibliography}
\end{frame}

\section*{Extra Slides}
\begin{frame}[fragile]{Weighted Rank Order Correlation Coefficient}
\only<+->{
We want to quantify the similarity between two rank orders of dendrites $R$ and $Q$, which both have length $n$.}

\only<+->{
Standard approach: Spearman's rank order coefficient 
$$
\rho(R,Q) = 1 - \frac{6\sum_i^nD_i^2}{n(n	^2-1)}
$$
where $D_i=R_i-Q_i$.}

\only<+->{
Problem: Many of the dendrites are not significant for classification but affect $\rho$ greatly. We would like to be able to \textbf{weight our} $\mathbf{D_i}$'s.}
\end{frame}

\begin{frame}[fragile]{Weighted Rank Order Correlation Coefficient}
\only<+->{
We want our rank order coefficient to be an affine linear function of $\sum_iw_iD_i^2$:

$$
\rho_W(R, Q) = A+B\sum_i^nw_iD_i^2
$$}
\only<+->{
If $R=Q$, we want $\rho_W=1$, and if the rankings are reversed, we want $\rho_W=-1$.}

\only<+->{
The first condition gives us $A=1$, and the second one
$$
B=\frac{-2}{\sum_i^nw_i(n-2i+1)^2} 
$$}
\only<+->{
Thus:
$$
\boxed{\rho_W(R,Q) = 1-\frac{-2\sum_i^nw_iD_i^2}{\sum_i^nw_i(n-2i+1)^2}}
$$}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\only<+->{
We would like to visualize the time evolution of the population response in a coordinate system in which the axes represent stimulus strength and choice.}

\only<+->{
To that end we use linear regression to write the normaized response of dendrite $i$ at time $t$ in trial $k$ as a linear combination of these task variables:
$$
r_k^{i,t} = \beta_1^{i,t}choice_k+\beta_2^{i,t}stimulus_k+\beta_3^{i,t}
$$}

\only<+->{
The regression coefficients $\beta_{\nu}^{i,t}$ describe how much the activity of dendrite $i$ at time $t$ in trial $k$ corresponds with variable $\nu$.}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
We define
$$
\mathbf{F} = 
\begin{bmatrix}
choice_1 & ... & choice_n \\
stimulus_1 & ... & stimulus_n \\
1 & ... & 1
\end{bmatrix}
$$
and estimate for each dendrite $i$ and timepoint $t$
$$
\boldsymbol{\beta^{i,t}} = \mathbf{(FF^T)^{-1}Fr^{i,t}}
$$
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\only<+->{
In total we have 14 conditions: 7 stimuli X two choices.}

\only<+->{
We define:
$$
\mathbf{x^{c,t}}
$$
as the (trial-averaged) population response to condition $c$ (for example lick at stimulus 1.00) at time $t$, which is a vector of $N_{unit}$ length.}

\only<+->{
The goal is to use $\boldsymbol{\beta}$ to find a two-dimensional subspace of the dendrite space into which we can transform $\mathbf{x^{c,t}}$.}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\only<+->{
We then use PCA to denoise the data matrix
$$
\mathbf{X}=
\begin{bmatrix}
 &  &  & \\
\mathbf{x_{c_1,t_1}} & ... & \mathbf{x_{c_1,t_n}} & ... & \mathbf{x_{c_m,t_n}}
\\
  &  &  &
\end{bmatrix}
$$}

\only<+->{
and we apply the denoising matrix to all $\boldsymbol{\beta^{\nu, t}}$ as well, which yields the denoised regression vectors
$$
\boldsymbol{\beta_{pca}^{\nu,t}}
$$}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\only<+->{
Since the $\boldsymbol{\beta_{pca}^{\nu,t}}$ are time variing but we would like a coordinate system that is fixed in time, we "freeze" them at the point in time at which there is maximum correlation:}
\only<+->{
\[
\boldsymbol{\beta^{\nu}_{max}}=\boldsymbol{\beta_{pca}^{\nu,t^{\nu}_{max}}}
\]
\[
t^{\nu}_{max} = argmax_t||\boldsymbol{\beta_{pca}^{\nu,t}}||
\]}
\end{frame}

\begin{frame}[fragile]{Population Response in Task Variable Space}
\only<+->{
We would like these $\boldsymbol{\beta^{\nu}_{max}}$ to be the basis vectors of our new coordinate system, however, they are not yet orthogonal.} 
\only<+->{We fix this by applying QR-decomposition to 
$$
\mathbf{B^{max}}=
\begin{bmatrix}
\mathbf{\beta^1_{max}} & \mathbf{\beta^2_{max}}
\end{bmatrix} = \mathbf{QR}
$$
where $\mathbf{Q}$ is an orthogonal matrix whose columns $\boldsymbol{\beta^{\nu}_{\bot}}$ are the \textbf{basis vectors} of our new coordinate system. We can now transform our data into it.}
\end{frame}
